R Notebook

### 2020-01-16 src/comb-maxent.Rmd
update to combined maxent

---
title: "maxent model avec xvalidation (year-based test training)"
author: "Samantha Andrews"
output: html_notebook
---

# Overview
A note to anyone who might happen to stumble across this... I am a beginner in R and have had no exposure to similar languages. I don't know what I'm doing. The code herein is unlikely to be elegant and there area probably more efficient ways of running the code.

Built with 'r getRversion()'..

# Package dependencies
You can install and load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). Note this function checks to see if the packages are installed first before loading.

@@

#load the presence-background dataset 
("../output/bio/presab_nafo_yr.csv")

@@

#monthly model data selection and prep
For each monthly model, subset the data to the month you want, and the variables you want (Based on spearmans and VIF)

```{r}
janvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_winter")
marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")
decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface", "nao_sample", "nao_prev", "nao_winter", "amo_sample", "amo_winter")

mthnavar_lst <- list(jan = janvar, mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
```

select month
```{r month of interest -change}
mthna <- "jan"
mthno <- "1"
```

@@

normal model..
mod12 <- maxent(x = prab, p = pa, file = .asc, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE")) 
mod12

```{r train and test inc AIC for reg val}
presyrs <-  #this is to get years with complete presence data

modelmonth <- cbind(paste0(mthna ,unique(prback$year)))
aicetc <- read.csv("../data/env/AICcCalculations.csv", header = TRUE) #pre-created csv file
betavallist <- cbind(aicetc$regularization)

for (year in 1:nrow(modelmonth)){
  timeslice <- modelmonth[year] #testing timeslice
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(prback, year != yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  test <-  subset(prback, year == yr) #creates the testing dataset
  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  for (h in 1:nrow(betavallist)){
    betaval <-  betavallist[h]
    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c('plots=FALSE', "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betavallist[h])))

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2])) #dodgy 
    bic <- (bick*log(sum(trainpa == 1)) - 2*11) #dodgy

    for (i in 1:nrow(aicetc)){
    if(aicetc$regularization[i] == betaval){
      aicetc$train_AUC[i] <- trainauc
      aicetc$no_trainng_points[i] <- notrainingpoints
      aicetc$coefficients[i] <- nocoeff
      aicetc$sum.of.logs[i] <- sumsamplepred
      aicetc$test_AUC[i] <- evauc
      aicetc$test_evcor[i] <- evcor
      aicetc$test_evpcor[i] <- evpcor
      aicetc$test_nopresence[i] <- evnp
      aicetc$test_nobackground[i] <- evna
      aicetc$test_maxsumsenspec[i] <- evmaxtprtnr
      aicetc$test_maxkap[i] <- evmaxkap
      aicetc$test_no_omission[i] <- evno_omission
      aicetc$test_prevalence[i] <- evprevalence
      aicetc$test_equal_sens_spec[i] <- evequal_sens_spec
      aicetc$test_sensitivity[i] <- evsensitivity
      aicetc$test_bic[i] <- bic #not sure this is correct...
      aicetc$beta[i] <- betaval
      #calculate minimal predicted area using ecospat
      aicetc$mpa_0.9_cum[i] <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%

    aicetc$AIC[i] <- ((2*aicetc$coefficients[i])-(2*aicetc$sum.of.logs[i]))
    aicetc$AICc[i] <- (-2*(aicetc$sum.of.logs[i])+((2*aicetc$coefficients[i])*(aicetc$no_trainng_points[i]/(aicetc$no_trainng_points[i] - aicetc$coefficients[i]-1))))
    }
   }
  }

write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

}

#model selection
next step... select the models you want for each year and run those models again (primarily based on lowest AICc but check test AUC and other things too)

you need to go through each aic file manually....

KEEP A NOTE OF THESE IN ONENOTE

#run final model
```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

@@

  if (yr == "1998"){
    betaval <- 0.25
  } else if (yr == "2001"){
    betaval <- 0.75
  } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
    betaval <- 1
  } else if (yr == "2005"){
    betaval <- 1.25
  } else if (yr == "1999" | yr == "2000"| yr == "2009"){
    betaval <- 1.5
  } else if (yr == "2007"){
    betaval <- 1.75
  } else if (yr == "2010"){
    betaval <- 2
  } else if (yr == "2008"){
    betaval <- 2.25
  }

    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))

@@

    #continuous boyce index
    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
    predPres <- predPres$Cloglog.prediction
    predBg <- predBg$Cloglog
    cbi <- contBoyce(pres = predPres, bg = predBg)

@@

    write.csv(aicetc, paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
    stackdeplst[[year]] <- stackdep
    write.csv(stackdep, paste0("../output/maxent/models/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

   for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_prev <- raster(unlayer_yr$nao_prev[td])
      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_prev) <- "nao_prev"
      nao_prev <- projectRaster(nao_prev, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      amo_sample <- raster(unlayer_yr$amo_sample[td])
      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(amo_sample) <- "amo_sample"
      amo_sample <- projectRaster(amo_sample, aea)
      amo_winter <- raster(unlayer_yr$amo_winter[td])
      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(amo_winter) <- "amo_winter"
      amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)

      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
      predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png

      #mess

      #messev <- dismo::mess(x = stack, v = mess_pr, full = FALSE)
      #messplot <- plot(messev, png(paste0("../output/maxent/predictions/", mthna, "/", "mess_", stackn, ".png")))
      #dev.off()
      # not working how i'd like....
    }
}

stdep <- do.call(rbind, stackdeplst)
write.csv(stdep, paste0("../output/maxent/models/", timeslice, "/stack_depth.csv"), row.names = FALSE)
```

```{r avg maxentReg}

for (i in 1:length(modelmonth)){
  timeslice <- modelmonth[i]
  yr <- str_sub(timeslice, -4) #the year being used for testing
  if(yr == 1998){
    aic1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 1999){
    aic1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2000){
    aic2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2001){
    aic2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2002){
    aic2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2003){
    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2004){
    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2005){
    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2006){
    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2007){
    aic2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2008){
    aic2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2009){
    aic2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2010){
    aic2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2011){
    aic2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2012){
    aic2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2013){
    aic2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2014){
    aic2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2015){
    aic2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
  }
}

aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)

aicm <- colMeans(aic)
aicm$mod <- "avg"
avgAIC <- rbind(aic, aicm)

maxr$Species <- 1
maxrm <- colMeans(maxr)
maxrm$Species <- "avg"
maxrm <- rbind(maxr, maxrm)

write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
```

summary table of average maxent results for each month
```{r average AIC etc for each month in one table}
aic_lst <- list.files("../output/maxent/models/", pattern = "avgAIC.csv", full.names = TRUE)
aicsum <- data.frame(matrix(ncol = 11, nrow = length(aic_lst)))
colnames(aicsum) <- c("Month", "BetaVal", "AICc", "Test AUC", "Training AUC", "Correlation", "Correlation P-Val", "TSS", "CBI", "No Omissions", "Prevalence")

for (a in 1:length(aic_lst)){
  aicr <- read.csv(aic_lst[a])
  mth <- str_sub(aic_lst[a], start = 25, end = 27) #selects the month from the filename
  aicsum$Month[a] <- mth
  aicres <- tail(aicr, 1)
  aicsum$BetaVal[a] <- aicres$beta
  aicsum$AICc[a] <- aicres$AICc
  aicsum$`Test AUC`[a] <- aicres$test_AUC
  aicsum$`Training AUC`[a] <- aicres$train_AUC
  aicsum$Correlation[a] <- aicres$test_evcor
  aicsum$`Correlation P-Val` <- aicres$test_evpcor
  aicsum$TSS[a] <- aicres$tss
  aicsum$CBI[a] <- aicres$cbi
  aicsum$`No Omissions`[a] <- aicres$test_no_omission
  aicsum$Prevalence[a] <- aicres$test_prevalence
}

mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
aicsum <- aicsum[match(mthord, aicsum$Month), ] #reorder based on mthord
write.csv(aicsum, "../output/maxent/models/aicsummary.csv", row.names = FALSE)

stacked permutation importance plot on average

```{r permutation & jack data}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
perm <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(perm) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
jack <- data.frame(matrix(ncol = 10, nrow = length(maxr_lst)))
colnames(jack) <- c("Month", "amo_sample", "amo_winter", "chl_surface", "nao_prev", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")

for (m in 1:length(maxr_lst)){
  maxr <- read.csv(maxr_lst[m])
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  perm$Month[m] <- mth
  jack$Month[m] <- mth
  maxt <- tail(maxr, 1)
  perm$amo_sample[m] <- maxt$amo_sample.permutation.importance
  perm$amo_winter[m] <- maxt$amo_winter.permutation.importance
  perm$chl_surface[m] <- maxt$chl_surface.permutation.importance
  perm$nao_prev[m] <- maxt$nao_prev.permutation.importance
  perm$nao_sample[m] <- maxt$nao_sample.permutation.importance
  perm$nao_winter[m] <- maxt$nao_winter.permutation.importance
  perm$o2_depth[m] <- maxt$o2_depth.permutation.importance
  perm$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
  perm$temp_depth[m] <- maxt$temp_depth.permutation.importance

  jack$amo_sample[m] <- maxt$Training.gain.without.amo_sample
  jack$amo_winter[m] <- maxt$Training.gain.without.amo_winter
  jack$chl_surface[m] <- maxt$Training.gain.without.chl_surface
  jack$nao_prev[m] <- maxt$Training.gain.without.nao_prev
  jack$nao_sample[m] <- maxt$Training.gain.without.nao_sample
  jack$nao_winter[m] <- maxt$Training.gain.without.nao_winter
  jack$o2_depth[m] <- maxt$Training.gain.without.o2_depth
  jack$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
  jack$temp_depth[m] <- maxt$Training.gain.without.temp_depth
}

# now for january

maxr_lstj <- list.files("../output/maxent/models/", pattern = "janavgmaxentresults.csv", full.names = TRUE)
permj <- data.frame(matrix(ncol = 7, nrow = length(maxr_lstj)))
colnames(permj) <- c("Month", "chl_surface", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")
jackj <- data.frame(matrix(ncol = 7, nrow = length(maxr_lstj)))
colnames(jackj) <- c("Month", "chl_surface", "nao_sample", "nao_winter", "o2_depth", "salinity_depth", "temp_depth")

for (m in 1:length(maxr_lstj)){
  maxr <- read.csv(maxr_lstj[m])
  mth <- str_sub(maxr_lstj[m], start = 25, end = 27) #selects the month from the filename
  permj$Month[m] <- mth
  jackj$Month[m] <- mth
  maxt <- tail(maxr, 1)
  permj$chl_surface[m] <- maxt$chl_surface.permutation.importance
  permj$nao_sample[m] <- maxt$nao_sample.permutation.importance
  permj$nao_winter[m] <- maxt$nao_winter.permutation.importance
  permj$o2_depth[m] <- maxt$o2_depth.permutation.importance
  permj$salinity_depth[m] <- maxt$salinity_depth.permutation.importance
  permj$temp_depth[m] <- maxt$temp_depth.permutation.importance

  jackj$chl_surface[m] <- maxt$Training.gain.without.chl_surface
  jackj$nao_sample[m] <- maxt$Training.gain.without.nao_sample
  jackj$nao_winter[m] <- maxt$Training.gain.without.nao_winter
  jackj$o2_depth[m] <- maxt$Training.gain.without.o2_depth
  jackj$salinity_depth[m] <- maxt$Training.gain.without.salinity_depth
  jackj$temp_depth[m] <- maxt$Training.gain.without.temp_depth
}

permu <- rbind.fill(perm, permj)
jackk <- rbind.fill(jack, jackj)

mthord <- c("jan", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by
permu <- permu[match(mthord, permu$Month), ] #reorder based on mthord
jackk <- jackk[match(mthord, jackk$Month), ] #reorder based on mthord

write.csv(permu, "../output/maxent/models/permutation_avg.csv", row.names = FALSE)
write.csv(jackk, "../output/maxent/models/jack_avg.csv", row.names = FALSE)

```{r}
maxr_lst <- list.files("../output/maxent/models/", pattern = "maxentresults.csv", full.names = TRUE)
maxr_lst <- maxr_lst[-4] #removing january because this month has slightly different variables than the others
jack_df <- list()
maxr_lstj <- list.files("../output/maxent/models/", pattern = "janavgmaxentresults.csv", full.names = TRUE)
jack_dfj <- list()

for (m in 1:length(maxr_lst)){
  res <- read.csv(maxr_lst[m], header = TRUE)
  mth <- str_sub(maxr_lst[m], start = 25, end = 27) #selects the month from the filename
  res_df <- tail(res, 1)
  var_df <- res_df[,c(grep("Training.gain.with*",names(res_df)))]
  var_df$id <- 1:nrow(var_df)
  var_df <- melt(var_df, id.vars=c("id"))
  jack_with <- var_df[grep("with.only",var_df$variable),]
  jack_with$Type_of_Run <- "With only the variable"
  jack_without <- var_df[grep("without",var_df$variable),]
  jack_without$Type_of_Run <- "Without the variable"
  gain_all <- res_df[c(grep("Regularized.",names(res_df)))]
  gain_all$id <- 1:nrow(gain_all)
  gain_all <- melt(gain_all, id.vars=c("id"))
  gain_all$Type_of_Run <- "With all variables"
  gain_all$variable <- "all_variables"
  jack_table <- rbind(jack_with,jack_without,gain_all)
  jack_table$state <- mth

>>>>

@@

load the presence-background dataset 
("../output/bio/presab_nafo_yr.csv")

select month
```{r month of interest -change}
mthna <- "mar"
mthno <- "3"
```

@@

monthly model data selection and prep
For each monthly model, subset the data to the month you want, and the variables you want (Based on spearmans and VIF)

```{r}
marvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
aprvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
mayvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
junvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
julvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
augvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
sepvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
octvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
novvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")
decvar <- c("occurrence", "year", "temp_depth", "salinity_depth", "o2_depth", "chl_surface")

mthnavar_lst <- list(mar = marvar, apr = aprvar, may = mayvar, jun = junvar, jul = julvar, aug = augvar, sep = sepvar, oct = octvar, nov = novvar, dec = decvar)
```

@@

```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/test-comb/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

@@

  if (yr == "2003"){
    betaval <- 1
  } else if (yr == "1999" | yr == "2002" | yr == "2008" | yr == "2009") {
    betaval <- 1.25
  } else if (yr == "2005" | yr == "2006"){
    betaval <- 1.5
  } else if (yr == "2000"){
    betaval <- 1.75
  } else if (yr == "2001" | yr == "2007"){
    betaval <- 2.25
  } else if (yr == "2010"){
    betaval <- 3.5
  }

    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))

    saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))

@@

    #continuous boyce index
    predPres <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    predBg <- read.csv(paste0("../output/maxent/models/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
    predPres <- predPres$Cloglog.prediction
    predBg <- predBg$Cloglog
    source("../src/contBoyce.R")
    cbi <- contBoyce(pres = predPres, bg = predBg)

@@

    write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

    mod <- readRDS(paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))

    yr <- as.numeric(yr)
 unlayer_yr <- subset(unlayer, ascyears == yr)
    #unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname))
    stackdeplst[[year]] <- stackdep
    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

  for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      #temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_depth[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      #chl_depth <- projectRaster(chl_depth, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      #o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      #salinity_depth <- projectRaster(salinity_depth, aea)
      # nao_sample <- raster(unlayer_yr$nao_sample[td])
      # crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      # names(nao_sample) <- "nao_sample"
      # nao_sample <- projectRaster(nao_sample, aea)
      # nao_prev <- raster(unlayer_yr$nao_prev[td])
      # crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      # names(nao_prev) <- "nao_prev"
      # nao_prev <- projectRaster(nao_prev, aea)
      # nao_winter <- raster(unlayer_yr$nao_winter[td])
      # crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      # names(nao_winter) <- "nao_winter"
      # nao_winter <- projectRaster(nao_winter, aea)
      # amo_sample <- raster(unlayer_yr$amo_sample[td])
      # crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      # names(amo_sample) <- "amo_sample"
      # amo_sample <- projectRaster(amo_sample, aea)
      # amo_winter <- raster(unlayer_yr$amo_winter[td])
      # crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      # names(amo_winter) <- "amo_winter"
      # amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface)

      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
      # predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png
  }
}
```
